\pkg{Rgbp}: An R Package for Gaussian, Poisson, and Binomial Hierarchical Modeling

### 2013-08-29 jss/article.tex
back
edit2

%%uncomment line below for double-spacing
\linespread{2}

@@

\subsection{Coverage Estimation Process}
Next, \code{coverage} fits the Normal-Normal model 500 times using the 500 pseudo-datasets in order to obtain 500$\times k$ interval estimates,  \{$(\hat{\mu}^{(i)}_{j, ~low}, ~\hat{\mu}^{(i)}_{j, ~upp}), ~j=1,\ldots, k,~ i=1, \ldots, 500$\}.  And then within each group, it will check simply how often the interval estimates include true parameters, or calculate Rao-Blackwellized coverage estimate.

\subsubsection{Simple Unbiased Coverage Estimates}
Let's define an indicator variable, $I^{(i)}_{j}$, which is 1 if the $j$-th group's interval estimate from the $i$-th pseudo-dataset includes $\mu^{(i)}_{j}$ and 0 otherwise. Then \code{coverage} estimates the coverage probability via the sample mean of these indicator variables for each group $j$, averaging over all the possible $\mu^{(i)}_{j}$ and $y^{(i)}_{j}$ for given $A$ and $\beta$. This provides $k$ simple unbiased coverage estimates. For example, $\frac{1}{500}\sum_{i=1}^{500}I^{(i)}_{1}$ is the simple unbiased coverage estimate for the first group ($j=1$), accounting for possible randomness in $\mu^{(i)}_{1}$ and $y^{(i)}_{1}$, $i=1, \ldots, 500,$ given a specific $(A, \beta)$. As the number of pseudo-datasets increases, this simple unbiased coverage estimate converges to the true coverage probability of the first group given $A$ and $\beta$.

\subsubsection{Rao-Blackwellized Unbiased Coverage Estimates}
Rao-Blackwellization improves accuracy of an unbiased estimator by taking a conditional expectation given a sufficient statistic. Based on this idea, we use $E(I^{(i)}_{j}\vert y^{(i)}_{j}, A, \beta$), where $A$ and $\beta$ are given and $y^{(i)}_{j}$ is the sufficient statistic for the $j$-th group in the $i$-th pseudo-dataset. This expectation is the same as Pr$(\hat{\mu}^{(i)}_{j, low}\le \mu^{(i)}_{j} \le\hat{\mu}^{(i)}_{j, upp}\vert y^{(i)}_{j}, A, \beta)$, where ($\hat{\mu}^{(i)}_{j, low}$, $\hat{\mu}^{(i)}_{j, upp}$) is the $j$-th group's interval estimate on the $i$-th dataset. We can calculate this probability  because we know the distribution of ($\mu^{(i)}_{j} \vert y^{(i)}_{j}, A, \beta$) in (\ref{normalpost}). Note that conditioning on $y^{(i)}_{j}$ is equivalent to conditioning on $\mathbf{y}^{(i)}_{(k\times1)}$ as long as $A$ and $\beta$ are known. Then, we estimate the first group's coverage probability by $\frac{1}{500}\sum_{i=1}^{500}E(I^{(i)}_{1}\vert y^{(i)}_{1}, A, \beta)$, which is also unbiased but with smaller variance than the previous simple estimator, $\frac{1}{500}\sum_{i=1}^{500}I^{(i)}_{1}$.

@@

For reference, \code{raw.resultRB} and \code{raw.resultS}, each row of which is about each hospital, in \code{pcv} contain all the individual simulation results.
% For reference, it took 10 seconds for MacBook Pro with a 2.3 GHz dual-core Intel i5 CPU to run 10,000 simulations.

@@

As for average simple unbiased estimates for each school, 
\begin{CodeChunk}
\begin{CodeInput}
R> gcv$coverageS
\end{CodeInput}
\begin{CodeOutput}
[1] 0.970 0.943 0.957 0.968 0.956 0.969 0.961 0.976
\end{CodeOutput}
\end{CodeChunk}

All the simulation results (0-1 coverage and RB coverage estimates) are saved in two $8\times1,000$ matrices, \code{gcv$raw.resultRB} and \code{gcv$raw.resultS}, each row of which is about each school.

@@

Now, let's see following shrinkage and 95\% interval plots for more intuition. It is obvious that sample means (empty dots) on the upper horizontal line shrink towards two different means, 0.233 and 0.310. For reference, the red line symbols on dots are for when two or more points have the same mean and are plotted over each other. For example, five players (from the 11th player to the 15th) have the same sample mean (0.222) and at this point on the upper horizontal line, there are short red lines toward five directions.

@@

In Figure \ref{fig:baseball} the 95\% interval plot shows us range of true batting average for each player, which clarifies the regression toward the mean (RTTM) within two groups. Let's see the 10th, 14th, and 15th players on the graph for example. They are outfielders but their observed batting averages are far lower than the first five outfielders. RTTM means that it can happen by their bad luck though in the long run their batting averages will come back to their expected ones as outfielders (blue horizontal lines). So, their posterior means (red dots) can be interpreted as their RTTM.

@@

And simple unbiased coverage estimates and their standard errors for each player are
\begin{CodeChunk}
\begin{CodeInput}
R> bcv$coverageS
\end{CodeInput}
\begin{CodeOutput}
 [1] 0.969 0.967 0.973 0.980 0.970 0.974 0.965 0.976 0.969 0.972 0.980 0.972 
[13] 0.957 0.969 0.972 0.982 0.978 0.984
\end{CodeOutput}
\end{CodeChunk}
\begin{CodeChunk}
\begin{CodeInput}
R> bcv$se.coverageS
\end{CodeInput}
\begin{CodeOutput}
 [1] 0.0045 0.0059 0.0066 0.0049 0.0052 0.0044 0.0042 0.0054 0.0056 0.0053 0.0049 
[11] 0.0056 0.0042 0.0049 0.0056 0.0057 0.0052 0.0052
\end{CodeOutput}
\end{CodeChunk}

All the simulation results are saved in two $18\times1,000$ matrices, \code{bcv$raw.resultRB} and \code{bcv$raw.resultS}, each row of which is about each player.

>>>>

%%uncomment line below for double-spacing
%\linespread{1}

@@

\subsection{Coverage Estimation Process}
Next, \code{coverage} fits the Normal-Normal model 500 times using the 500 pseudo-datasets in order to obtain 500$\times k$ interval estimates,  \{$(\hat{\mu}^{(i)}_{j, ~low}, ~\hat{\mu}^{(i)}_{j, ~upp}), ~j=1,\ldots, k,~ i=1, \ldots, 500$\}.  Here, we introduce both simple and Rao-Blackwellized unbiased coverage estimators but \code{coverage} displays only Rao-Blackwellized one.

\subsubsection{Simple Unbiased Coverage Estimates}
Let's define an indicator variable, $I^{(i)}_{j}$, which is 1 if the $j$-th group's interval estimate from the $i$-th pseudo-dataset includes $\mu^{(i)}_{j}$ and 0 otherwise. Then \code{coverage} estimates the coverage probability via the sample mean of these indicator variables for each group $j$, averaging over all the possible $\mu^{(i)}_{j}$ and $y^{(i)}_{j}$ for given $A$ and $\beta$. 
\begin{equation}
\bar{I}_{j}\equiv \frac{1}{N_{sim}}\sum_{i=1}^{N_{sim}}I^{(i)}_{j},~ j=1, 2, \ldots, k.
\end{equation}

This provides $k$ simple unbiased coverage estimates. For example, $\frac{1}{500}\sum_{i=1}^{500}I^{(i)}_{1}$ is the simple unbiased coverage estimate for the first group ($j=1$), accounting for possible randomness in $\mu^{(i)}_{1}$ and $y^{(i)}_{1}$, $i=1, \ldots, 500,$ given specific values of $A$ and $\beta$.

Each indicator has the independent and identically distributed (\emph{iid}) Bernoulli($p_{cov, j}$) distribution, where $p_{cov, j}$ is the true coverage probability for the $j$-th group depending on specific values of $A$ and $\beta$.

So, its unbiased variance estimator  for $Var(\bar{I}_{j})$ is simply the sample variance of indicators over the number of simulations.
\begin{equation}
\widehat{Var}(\bar{I}_{j})\equiv\frac{1}{N_{sim}(N_{sim}-1)}\sum_{i=1}^{N_{sim}}(I^{(i)}_{j}-\bar{I}_{j})^{2},~ j=1, 2, \ldots, k.
\end{equation}

\subsubsection{Rao-Blackwellized Unbiased Coverage Estimates}
Rao-Blackwellization improves accuracy of an unbiased estimator by taking more information (sufficient statistic) into consideration. Based on this idea, we use $E(I^{(i)}_{j}\vert y^{(i)}_{j}, A, \beta$), where $A$ and $\beta$ are given and $y^{(i)}_{j}$ is the sufficient statistic for the $j$-th group in the $i$-th pseudo-dataset. 
\begin{equation}\label{RB}
\bar{I}_{RB, j}\equiv \frac{1}{N_{sim}}\sum_{i=1}^{N_{sim}}E(I^{(i)}_{j}\vert y^{(i)}_{j}, A, \beta),~ j=1, 2, \ldots, k.
\end{equation}

The expectation in (\ref{RB}) is the same as Pr$(\hat{\mu}^{(i)}_{j, low}\le \mu^{(i)}_{j} \le\hat{\mu}^{(i)}_{j, upp}\vert y^{(i)}_{j}, A, \beta)$, where ($\hat{\mu}^{(i)}_{j, low}$, $\hat{\mu}^{(i)}_{j, upp}$) is the $j$-th group's interval estimate on the $i$-th dataset. We can calculate this probability  because we know the distribution of ($\mu^{(i)}_{j} \vert y^{(i)}_{j}, A, \beta$) in (\ref{normalpost}). Note that conditioning on $y^{(i)}_{j}$ is equivalent to conditioning on $\mathbf{y}^{(i)}_{(k\times1)}$ as long as $A$ and $\beta$ are known.

For example, we estimate the first group's coverage probability (that depend on specific $A$ and $\beta$) by $\frac{1}{500}\sum_{i=1}^{500}E(I^{(i)}_{1}\vert y^{(i)}_{1}, A, \beta)$, which is also unbiased for $p_{cov, 1}$ but with smaller variance than the previous simple estimator, $\frac{1}{500}\sum_{i=1}^{500}I^{(i)}_{1}$.

Under the setting where one dataset ($\mathbf{y}^{(i)}_{(k\times1)}$) is simulated per one true parameter set ({\boldmath $\mu$}$^{(i)}_{(k\times1)})$, the unbiased variance estimator for Var($\bar{I}_{RB, j}$) follows.
\begin{equation}
\widehat{Var}(\bar{I}_{RB, j})\equiv\frac{1}{N_{sim}(N_{sim}-1)}\sum_{i=1}^{N_{sim}}\bigg(E(I^{(i)}_{j}\vert y^{(i)}_{j}, A, \beta)-\bar{I}_{RB, j}\bigg)^{2},~ j=1, 2, \ldots, k.
\end{equation}

@@

For reference, two $31 \times 1,000$ matrices \code{raw.resultRB} and \code{raw.resultS}, each row of which is about each hospital, in \code{pcv} contain all the individual estimates.
% For reference, it took 10 seconds for MacBook Pro with a 2.3 GHz dual-core Intel i5 CPU to run 10,000 simulations.

@@

All the individual RB coverage estimates are saved in the $8\times1,000$ matrix, \code{gcv$raw.resultRB}, each row of which is about each school.

@@

Now, let's see the shrinkage and 95\% interval plots for more intuition in Figure \ref{fig:baseball}. It is obvious that sample means (empty dots) on the upper horizontal line shrink towards two different means, 0.233 and 0.310. For reference, the red line symbols on dots are for when two or more points have the same mean and are plotted over each other. For example, five players (from the 11th player to the 15th) have the same sample mean (0.222) and at this point on the upper horizontal line, there are short red lines toward five directions.

@@

The 95\% interval plot shows us range of true batting average for each player, which clarifies the regression toward the mean (RTTM) within two groups. Let's see the 10th, 14th, and 15th players on the graph for example. They are outfielders but their observed batting averages are far lower than the first five outfielders. RTTM means that it can happen by their bad luck though in the long run their batting averages will come back to their expected ones as outfielders (blue horizontal lines). So, their posterior means (red dots) can be interpreted as their RTTM.

@@

All the simulation results are saved in the $18\times1,000$ matrix, \code{bcv$raw.resultRB}, each row of which is for each player.
